Efficient ancestry and mutation simulation with msprime 1.0

### 2021-08-15 paper.tex
spell check

The \msprime\ simulator~\citep{kelleher2016efficient,kelleher2020coalescent}
has greatly increased the scope of coalescent simulations,
and it is now straightforward to simulate millions of whole chromosomes.
The ``succinct tree sequence'' data
structure~\citep{kelleher2016efficient,kelleher2018efficient,kelleher2019inferring,
wohns2021unified},
originally introduced as part of \msprime, makes it possible to store
such large simulations in a few gigabytes, several orders
of magnitude smaller than commonly used formats.
The succinct tree sequence has also led to major advances in forwards-time
simulation~\citep{kelleher2018efficient,haller2018tree},
ancestry inference~\citep{kelleher2019inferring,wohns2021unified}
and calculation of population genetic
statistics~\citep{kelleher2016efficient,ralph2020efficiently}.
Through a rigorous open-source community development process,
\msprime\ has gained a large number of features since its introduction,
making it a highly efficient and flexible platform for population
genetic simulation.
This paper marks the release of \msprime\ 1.0.
We provide an overview of its extensive features,
demonstrate its performance advantages over alternative software,
and discuss opportunites for ongoing
open-source community-based development.

@@

\begin{figure}
\begin{center}
\includegraphics{figures/sweeps-perf}
\end{center}
\caption{\label{fig-selection-perf}Comparison of selective sweep simulation
performance in \msprime\ and \discoal\ (Intel Xeon E5-2680 CPU).
We report the total CPU time and maximum memory usage when simulating 100
replicates for 10 samples in a model
with a single selective sweep in its history where the benefical
allele had a scaled selection coefficient of $2Ns=1000$,
a per-base recombination rate of $10^{-9}$ and sequence
length varying from 1kb--200kb.}
\end{figure}

>>>>

The \msprime\ simulator~\citep{kelleher2016efficient,kelleher2020coalescent}
has greatly increased the scope of coalescent simulations,
and it is now straightforward to simulate millions of whole chromosomes.
The ``succinct tree sequence'' data
structure~\citep{kelleher2016efficient,kelleher2018efficient,kelleher2019inferring,
wohns2021unified},
originally introduced as part of \msprime, makes it possible to store
such large simulations in a few gigabytes, several orders
of magnitude smaller than commonly used formats.
The succinct tree sequence has also led to major advances in forwards-time
simulation~\citep{kelleher2018efficient,haller2018tree},
ancestry inference~\citep{kelleher2019inferring,wohns2021unified}
and calculation of population genetic
statistics~\citep{kelleher2016efficient,ralph2020efficiently}.
Through a rigorous open-source community development process,
\msprime\ has gained a large number of features since its introduction,
making it a highly efficient and flexible platform for population
genetic simulation.
This paper marks the release of \msprime\ 1.0.
We provide an overview of its extensive features,
demonstrate its performance advantages over alternative software,
and discuss opportunities for ongoing
open-source community-based development.

@@

\begin{figure}
\begin{center}
\includegraphics{figures/sweeps-perf}
\end{center}
\caption{\label{fig-selection-perf}Comparison of selective sweep simulation
performance in \msprime\ and \discoal\ (Intel Xeon E5-2680 CPU).
We report the total CPU time and maximum memory usage when simulating 100
replicates for 10 samples in a model
with a single selective sweep in its history where the beneficial
allele had a scaled selection coefficient of $2Ns=1000$,
a per-base recombination rate of $10^{-9}$ and sequence
length varying from 1kb--200kb.}
\end{figure}
